---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

options(knitr.duplicate.label = "allow")
```

# agroclimR: Streamlining Agroclimatic Data for Crop Modeling

<!-- badges: start -->
<!-- badges: end -->

## Overview

`agroclimR` is an R package developed to  facilitate the management and transformation of agroclimatic data into crop model-ready formats. It supports major modeling systems such as [DSSAT v 4.8](https://dssat.net/), [ORYZA v3.0](https://www.irri.org/oryza), and [Aquacrop v6.0](https://www.fao.org/aquacrop/). This toolkit serves as a resource for researchers and students specializing in agricultural sciences.

## Features

- **Efficient Data Management:** Provides a relational database structure that simplifies data handling and seamlessly integrates with the agroclimR framework. Learn more about the data structure in [data-format](https://jrodriguez88.github.io/agroclimR/articles/Estructura_de_Datos_para_Modelacion_de_Cultivos_con_agroclimR.html)
- **Easy Data Transformation to Crop Model-Ready Formats:** Utilizes R to convert various model format files efficiently.
- **Tidyverse Integration:** Employs tidyverse functions for smooth data manipulation and offers insightful data visualization capabilities using ggplot2.


## Installation

You can install the development version of agroclimR from [GitHub](https://github.com/jrodriguez88/agroclimR) with:

``` r
# install.packages("devtools")
devtools::install_github("jrodriguez88/agroclimR")
```

## Example: Creating Weather and Soil Model Format Files

### Weather Model Files

Here's a basic example demonstrating how to create a crop model weather file using agroclimR:

```{r example}
library(agroclimR)

path = "."
wth_data = weather
lat = 3.8
lon = -76.5
elev = 650

## basic example code to create ORYZA weather file
wth_files_oryza <- write_wth_oryza(path, "wth_oryza", wth_data, lat, lon, elev, tag = TRUE)
readLines(wth_files_oryza[1], n = 24) |> writeLines()
file.remove(wth_files_oryza)

## basic example code to create DSSAT weather file
wth_files_dssat <- write_wth_dssat(path, "wth_dssat", wth_data, lat, lon, elev)
readLines(wth_files_dssat[1], n = 10) |> writeLines()
file.remove(wth_files_dssat)
 
## basic example code to create AquaCrop weather file  
wth_files_aquacrop <- write_wth_aquacrop(path, "wth_aquacrop", wth_data, lat, lon, elev)
readLines(wth_files_aquacrop[1], n = 10) |> writeLines()
readLines(wth_files_aquacrop[2], n = 13) |> writeLines()
file.remove(wth_files_aquacrop)
 
```

### Soil Model Files

This is a basic example which shows you how to create a crop model Soil file with agroclimR:

```{r example2_soil}


soil_sample = dplyr::group_by(soil, NL) |> 
  dplyr::sample_n(1) |> dplyr::ungroup()


# basic example code to create ORYZA soil file
soil_files_oryza <- write_soil_oryza(id_name = "soil_oryza", soil_data = soil_sample)
readLines(soil_files_oryza[1], n = 27) |> writeLines()
file.remove(soil_files_oryza)

# basic example code to create DSSAT soil file
soil_files_dssat <- write_soil_dssat(path = ".", "soil_dssat", soil_sample)
readLines(soil_files_dssat[1], n = 15) |> writeLines()
file.remove(soil_files_dssat)

# basic example code to create Aquacrop soil file
soil_files_aquacrop <- write_soil_aquacrop(path = ".", "soil_dssat", soil_sample)
readLines(soil_files_aquacrop[1], n = 15) |> writeLines()
file.remove(soil_files_aquacrop)

 
```

## Connect

For inquiries or feedback, contact Jeferson Rodriguez-Espinoza at <jrodriguezespinoza@outlook.com>.

## Explore More

- [Documentation](https://jrodriguez88.github.io/agroclimR/)
- [DSSAT](https://dssat.net/)
- [ORYZA](https://www.irri.org/oryza)
- [Aquacrop](https://www.fao.org/aquacrop/)


