---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# agroclimR: A R-Toolkit for Agroclimatic Data Management and Crop Model File Creation

<!-- badges: start -->
<!-- badges: end -->

agroclimR is a package designed to streamline the management, manipulation, and generation of process-based model files for systems like [DSSAT v 4.8](https://dssat.net/), [ORYZA v3.0](https://www.irri.org/oryza), and [Aquacrop v6.0](https://www.fao.org/aquacrop/).

## Installation

You can install the development version of agroclimR from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("jrodriguez88/agroclimR")
```

## Example

This is a basic example which shows you how to create a crop model weather file with agroclimR:

```{r example}
library(agroclimR)

## basic example code

 write_wth_oryza(
   path = ".", id_name = "TEST", wth_data = weather,
   lat = 3.8, lon = -76.5, elev = 650)
 
 list.files(pattern = "TEST")
```



